TEMPERATURES AND METALLICITIES FOR 10,000 M DWARFS\\ IN THE \apogee\ SURVEY

### 2019-06-07 paper/draft.tex
change email
add ref

\correspondingauthor{Jessica Birky}
\email{jbirky@ucsd.edu}

@@

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations (e.g. \kepler, \tess), and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches, making detailed and precise knowledge of M dwarf chemical compositions key to constraining the formation scenarios and atmospheric conditions of potentially habitable exoplanets.

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade, such as the Sloan Digital Sky Survey (SDSS), and the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST) have dramatically increased the sample of known M dwarfs (\citealt{West:2011}, \citealt{Guo:2015}) and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level.
The launch of the Apache Point Observatory Galactic Evolution Experiment (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy \citep{Desphande:2013}, yet pipeline measurements \citep{Perez:2016} of $\teff$ and $\feh$ have only been determined precise to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K using atmospheric models \citep{Schmidt:2016}.

>>>>

\correspondingauthor{Jessica Birky}
\email{jbirky@uw.edu}

@@

Additionally, the low masses of M dwarfs make for easier detection of planets by radial velocity offset detections \citep{Trifonov:2018}, high planet-to-star radii ratios make for easier detection of exoplanet transits in light curve observations \citep{Nutzman:2008}, and shorter orbital periods allow for new planet discovery in less observation time than their higher mass counterparts. For these reasons, M dwarfs are the primary candidates for exoplanet detection searches (e.g. the NASA \kepler\ and Transiting Exoplanet Survey Satellite (\tess) missions), making detailed and precise knowledge of M dwarf chemical compositions key to constraining the formation scenarios and atmospheric conditions of potentially habitable exoplanets.

Advances in instrumentation with the lauch of several spectroscopic surveys in the past decade, such as the Sloan Digital Sky Survey (SDSS), and the Large Sky Area Multi-Object Fiber Spectroscopic Telescope (LAMOST) have dramatically increased the sample of known M dwarfs (\citealt{West:2011}, \citealt{Guo:2015} with catalogs of over 70,000 sources) and our potential to study M dwarf abundances on both exoplanet and galactic scales at a large statistical level.
The launch of the Apache Point Observatory Galactic Evolution Experiment (\apogee; \citealt{Majewski:2015}) survey as part of the Sloan Digital Sky Survey III mission \citep{Eisenstein:2011} has introduced the largest sample of M dwarfs observed with high resolution spectroscopy \citep{Desphande:2013}, yet pipeline measurements \citep{Perez:2016} of $\teff$ and $\feh$ have only been determined precise to 100\,K and 0.18\,dex down to $T_{\rm eff}\sim3550\,$K using atmospheric models \citep{Schmidt:2016}.
